Kepler and the Behemoth: Three Mini-Neptunes in a 40 Million Year Old Association

### 2022-09-05 papers/paper_review/response_1.txt

Sincerely,

Luke Bouma

******************************************
Reviewer #1 MODERATE CONCERNS

REVIEWER COMMENT
----------
Figure 1:
- It is interesting to note that all four planets are concentrated in X-Z, vb-vl*, and vl*-l. This could be because of the small fraction of space that the Kepler field surveyed. Is that accurate? Is there any way to indicated which stars were observed by Kepler in these figures?
- Note that in B&W only the + symbol has a different color than the rest, although since the symbols are different this wasn't much of an issue. The exception is that I can't tell which group is magenta and which lime green, so specifying which are the larger filled circles and which are the smaller would be helpful.
- Why is it important to show the projection of the solar velocity?
RESPONSE
----------
> - The Kepler field is shown in the top panel of the figure, between roughly l of 70 to 85, and b of 5 to 20.  This roughly corresponds to a cone in XYZ -- in case it helps intution, (l,b)=(90,0) corresponds to the line of X=0.  The resulting concentration of the planets in vl* we believe follows from the clustering in l, which is a direct result of the fraction of space that Kepler surveyed, combined with the solar velocity projection effect (described in point three below).
>
> - We have specified in the caption which are the larger and smaller filled circles.
>
> - We have briefly summarized in the first paragraph of section 2.2 why we have shown the projection of the solar velocity: it is because we are interested in how much of the spread in v_l* it might explain.  The short answer is "some, but not all of it".  To give a more complete explanation: the Sun moves at around 16 km/s toward (l,b)=(53deg,23deg), relative to a hypothetical circular orbit around the center of the galaxy.  This means that in our observed frame, nearby stars on average appear to move opposite this direction.  At l=53 and 53+180 degrees, the effect does not influence vl.  At l=53+90 degrees, the effect on vl (or equivalently vl*) is maximized, with amplitude = v_LSR = 16 km/s.  The minimum is at l=53+270 degrees.  Figure 11 of Zari, Brown et al 2018 provides some helpful intuition for how much of the variance in observed vl can be ascribed to this projection effect for stars near the galactic plane.  The lower-right panel of our Figure 1 however shows that while the LSR projection accounts for some of the change in observed v_l* in Cep-Her, the data and the gray band do not match!  There therefore needs to be some intrinsic differences in the UVW velocities of these stars (though v_r, v_phi, and v_z is perhaps the better coordinate system).

REVIEWER COMMENT
----------
If at all possible, please assure in Kerr+in prep that the nomenclature does not change such that CH-2 remains CH-2, or update this paper prior to publication.
RESPONSE
----------
> The sub-group identifier will remain the same.

REVIEWER COMMENT
----------
Line 248: I am skeptical of the few % false positive rate (later referenced as 1%). Once stars get hotter than K2, the young and field sequences overlap significantly and so field stars wouldn't likely be outliers in the CAMD. I'm also wondering if not all of the stars are shown on the CAMD; the text states 173 stars in RSG-5; it doesn't quite look to me like there are 173 stars on the upper left panel, but I could be wrong.
RESPONSE
----------
> Thank you for this comment -- it has helped us assess the false positive rates in RSG-5 more closely.
>
> In the color versus absolute magnitude diagram (CAMD), let us use (BP-RP)0 = 1.5 as the dividing line, redward of which stars can be easily seen to be on the pre-main-sequence (or at least elevated relative to typical field stars of the same color).  For RSG-5, there are 89 stars with (BP-RP)0 > 1.5.  One of them falls on the "field star" locus; the others fall on the pre-main-sequence locus.  This implies that, photometrically, 88/89= 98.9% of the candidate RSG-5 members redward of (BP-RP)0=1.5 are photometrically consistent with being on the pre-main-sequence.
>
> Regarding more massive stars, yes, there are plenty: see hr_rsg5.png attached to the response.  The caption of Figure 2 already explicitly states "The range of colors is truncated to emphasize the pre-main-sequence."  This visualization point is necessary because the full dynamic range in magnitude and color.  But yes, we cannot determine whether these stars earlier than ~K2 are young based on their location in the CAMD.  We can and do use stellar rotation.
>
> Using stellar rotation, we previously quoted in Section 2.3.2 that in RSG-5 41/47 stars have rotation periods consistent with being at least as fast as the Pleiades (87%).  Of the six inconsistent stars, two are the M-dwarfs previously discussed in the text.  The new point that emerged from our reanalysis concerns the other four stars without any detectable rotation period: three of them were the bluest three stars in our sample!  This implies that our blue cutoff was too lenient -- the amplitudes of spot-induced modulation plummet for young stars blueward of the Kraft break (Figure 10 in Rebull+2020, 2020AJ....159..273R).  We have propagated this through the manuscript: we now require (BP-RP)0 > 0.6, and G < 16.  The corresponding fraction of RSG-5 stars for which detected rotation periods is 39 detections / 42 expected, i.e. 93%.  
>
> The manuscript has been adjusted for this discussion, including the quantitative statements regarding the pre-main-sequence K and M dwarfs, as well as the rapidly-rotating FGK stars.

@@

REVIEWER COMMENT
----------
Line 122: It is worth noting that Zucker+2022 found that many of the "strings" from Kounkel & Covey are not likely real associations. Obviously, some of the associations K&C identify are real, but the extended groups that are more similar to Cep-Her may more often be artifacts.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 130, "identified three of the nearest subpopulations": subpopulations of Cep-Her?
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 140: Figure 1 reference seems to be specifically to the lower right panel, not both lower panels.
RESPONSE
----------
>

REVIEWER COMMENT
----------
*Line 162: vb and vl* not defined
RESPONSE
----------
> $v_b$ is the distance-corrected proper motion in the direction of increasing galactic latitude, and $v_{l^*}=v_l \cos b$ is the same in the direction of increasing galactic longitude after accounting for the local tangent plane correction.  We have added verbiage to this effect.

@@

REVIEWER COMMENT
----------
Line 392, citation to Chubak et al.: were these values determined following the methods of those authors, or by comparison to templates from that work?
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 394: Would be helpful to include a one-sentence description of Li measurement.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 413: Just wanted to confirm the 2015 date, presumably this was an archival spectrum from Kepler days.
RESPONSE
----------
> Confirmed.

REVIEWER COMMENT
----------
*Line 449: Teff values are only quoted for these two stars, and were not in previous sections, and the subsequent text also implies that for these stars, the spectra-based Teff values disagreed with the isochrone-based values. Either re-phrase, or discuss disagreement.
RESPONSE
----------
> Thank you for this wording catch -- the text has been updated to clarify that the temperatures as well as the other spectroscopic parameters agreed with the isochrone parameters within 1-sigma.  This is a rather brief way of hiding what in our view is more complicated than what needs to be discussed.  Table 1 states the adopted (spectroscopic) effective temperatures for all the stars.  We reported the spectroscopic effective temperatures in all cases because their formal uncertainties were smaller than those from the isochrone method, and they seemed less model-dependent.  The isochrone method yielded the following effective temperatures for the stars in Table 1:
> Kepler-1643: 4878 +/- 140 K
> KOI-7368: 5135 +/- 162 K
> KOI-7913 A: 4286 +/- 144 K
> KOI-7913 B 4006 +/- 335 K.
> As discussed in the second paragraph of Section 3, the quoted values are from MIST, while the uncertainties are a quadrature sum of a statistical and systematic component (where the latter is from the absolute difference between the PARSEC and MIST isochrones).  For KOI-7913 B, this systematic component is higher than for any other stars, because MIST gives Teff_MIST = 4005 +75 -73 K, while PARSEC gives Teff_PARSEC = 3679 +95 -60 K.  This is most likely due to different assumptions in the MIST and PARSEC models for how to treat young M-dwarfs.  The MIST models have known discrepancies between their synthetic photometry and observed open cluster data (e.g., Choi+2016 Section 8).  Possible explanations include starspots, or incomplete molecular line lists (e.g., Mann+2013, Rajpurohit+2013). The PARSEC isochrones might be a better choice for the lowest mass stars (KOI-7913A and KOI-7913B), since Chen et al (2014) implemented an empirical temperature-opacity calibration into the PARSEC models to create synthetic photometry that matches available data.  See for instance "hr_parsec_7913b.png" and "hr_mist_7913b.png" attached -- the 39.8 Myr MIST model diverges from the IC2602 and delta-Lyr loci at BP-RP>~1.6.  In comparison, the 39.8 Myr PARSEC model fits the observed locus down to BP-RP~=2.0.  
> Three additional plots are also included in the response: smemp_koi7913B.png, smemp_koi7913A.png, and smemp_kepler1643.png.  These plots give some diagnostic understanding of how SpecMatch-Empirical performs for these stars.  Note that a chi^2 of 5 corresponds to a reduced chi^2 of 1 in this model.  While the KOI-7913B spectrum is still the worst fit(!), our prefernce is for calibrating our reported effective temperatures to the HIRES spectral library, rather than to the isochronal method.  Any suggestions on possible methodological improvements would be appreciated.

### 2022-09-05 papers/paper/ms.tex
update response

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
  {\bf Positions and velocities of candidate members of the Cep-Her
  complex.}
  {\it Top row}: On-sky positions in galactic coordinates.  Black
  points are stars for which group membership is more secure than for
  gray points.  Kepler-1627 is in the outskirts of the $\delta$\ Lyr
  cluster \citep{bouma_kep1627_2022}, which is centered at $(l,b)
  \approx (66^\circ, 12^\circ)$.
  The Kepler footprint is shown in gray.
  {\it Middle row}: Galactic positions.  The Sun is at $(X, Y, Z) =
  (0, 0, 20.8)$\,pc; lines of constant heliocentric distance are
  shown between 250 and 400\,pc, spaced by 50\,pc.
  {\it Bottom row}: Galactic tangential velocities (left) and
  galactic longitudinal velocity versus galactic longitude (right).
  The gray band in the lower-right shows the $\pm$1-$\sigma$
  projection of the Solar velocity with respect to the local standard
  of rest \citep{schonrich_local_2010}.  There is a strong spatial and kinematic overlap between
  Kepler-1643 and RSG-5 (magenta).  The local population
  of candidate young stars around KOI-7368 and KOI-7913 is more
  diffuse -- we call this region ``CH-2'' (lime-green).
  The selection method for all of the stars 
  is described in Section~\ref{subsec:members}.
	\label{fig:XYZvtang}
	}
\end{figure*}

@@

The possibility that the $\delta$~Lyr cluster, RSG-5, and the
sub-populations identified by \citet{Kerr2021} share a common origin
has yet to be fully substantiated, but preliminary clustering results
from the 1\,kpc SPYGLASS analysis (R.~Kerr et al.\ in prep) suggest
the presence of contiguous stellar populations connecting each of
these groups in space-velocity coordinates.  
In other words, the stars appear to be comoving, though with a
continuous gradient in velocity as a function of position
(Figure~\ref{fig:XYZvtang}, lower panels).
In this work, our primary interest in the region stems
from the fact that a portion of it was observed by Kepler
(Figure~\ref{fig:XYZvtang}, top panel).  To further explore this
sub-population, we select candidate Cep-Her members through four
steps, the first three being identical to those described in Section~3
of \citet{Kerr2021}.  We briefly summarize them here.

@@

The CAMDs show that for RSG-5, all but one of the candidate
members are on a tight pre-main-sequence locus.  This implies a false
positive rate of a few percent, at most.  By comparison, our control
sample (the $\delta$~Lyr candidates) has a false positive rate of
$\approx$12\%, based on the number of stars that photometrically
overlap with the field population, while being outliers from the bulk
cluster population.  For CH-2, our membership selection gives 27
objects in the color range displayed, and 23 of them appear to be
consistent with being on the pre-main-sequence.  This implies a false
positive rate in CH-2 of $\approx$15\%.

@@

We selected stars suitable for gyrochronology by requiring $\bpmrpo
\geq 0.5$ to focus on FGKM stars that experience magnetic braking.
For TESS, we also restricted our sample to $G<16$, to ensure the stars
are bright enough to extract usable light curves from the full-frame
images.  The magnitude cut corresponds to $\bpmrpo < 2.6$
($\sim$M3V) at the relevant distances.  These cuts gave 19 stars in
CH-2 and 47 stars in RSG-5.  We extracted light curves from the TESS
images using the \texttt{unpopular} package \citep{hattorio_2021_cpm},
and regressed them against systematics with its causal pixel model.
We measured rotation periods using Lomb-Scargle periodograms and
visually vetted the results using an interactive program that allows
us to switch between TESS Cycles, select particular sectors, flag
stars with multiple periods, and correct half-period harmonics. For
ZTF, we used the same color cut to focus on FGKM stars, but restricted
the sample to $13 < G < 18$ to avoid the saturation limit on the
bright end and ensure sufficient photometric precision at the faint
end. We followed the procedure outlined in \citet{curtis_rup147_2020}:
we downloaded $8'\times8'$ image cutouts, ran aperture photometry for
the target and neighboring stars identified with Gaia, and used them
to define a systematics correction to refine the target light curves.

The lower panels of Figure~\ref{fig:age} show the results.  In RSG-5,
41/47 stars have rotation periods faster than the Pleiades (87\%).
This numerator omits the two stars with periods $>$$12$\,days visible
in the lower-left panel of Figure~\ref{fig:age}.  The age
interpretation for these latter stars, particularly the $\approx$M2.5
dwarf, is not obvious.  \citet{rebull_usco_2018} for instance have
found numerous M-dwarfs with 10-12 day rotation periods at ages of
USco ($\sim$$8$\,Myr), and some may still exist at ages of LCC
($\sim$$16$\,Myr; L.~Rebull submitted).  Regardless, since
nearly no field star outliers seem to be present on the RSG-5 CAMD,
the fact that we do not detect rotation periods for $\approx$13\% of
stars should perhaps be taken as an indication for the fraction of
stars for which rotation periods might not be detectable, due to
{e.g.}, pole-on stars having lower amplitude starspot modulation.

For CH-2, 13/19 stars have rotation periods that are obviously faster
than their counterparts in the Pleiades (68\%).  4 stars, not included
in the preceding numerator, are M-dwarfs with rotation periods between
10 and 12.5 days.  As previously noted, the age interpretation for these
M-dwarfs is not obvious.  Regardless, the $\approx$15\% false positive
rate for CH-2 determined from the CAMD seems consistent with our
fraction of detected rotation periods; RSG-5 was missing rotation
periods for $\approx$15\% of its candidate members, even though all of
its stars appear photometrically consistent with being on a single
pre-main-sequence locus.

>>>>

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
  {\bf Positions and velocities of candidate members of the Cep-Her
  complex.}
  {\it Top row}: On-sky positions in galactic coordinates.  Black
  points are stars for which group membership is more secure than for
  gray points.  Kepler-1627 is in the outskirts of the $\delta$\ Lyr
  cluster \citep{bouma_kep1627_2022}, which is centered at $(l,b)
  \approx (66^\circ, 12^\circ)$.
  The Kepler footprint is shown in gray.
  {\it Middle row}: Galactic positions.  The Sun is at $(X, Y, Z) =
  (0, 0, 20.8)$\,pc; lines of constant heliocentric distance are
  shown between 250 and 400\,pc, spaced by 50\,pc.
  {\it Bottom row}: Galactic tangential velocities (left) and
  galactic longitudinal velocity versus galactic longitude (right).
  The gray band in the lower-right shows the $\pm$1-$\sigma$
  projection of the Solar velocity with respect to the local standard
  of rest \citep{schonrich_local_2010}.  There is a strong spatial and kinematic overlap between
  Kepler-1643 and RSG-5 (magenta\added{; smaller circles}).  The local population
  of candidate young stars around KOI-7368 and KOI-7913 is more
  diffuse -- we call this region ``CH-2'' (lime-green\added{; larger
  circles}).
  The selection method for \replaced{all of the stars}{these groups}
  is described in Section~\ref{subsec:members}.
	\label{fig:XYZvtang}
	}
\end{figure*}

@@

The possibility that the $\delta$~Lyr cluster, RSG-5, and the
sub-populations identified by \citet{Kerr2021} share a common origin
has yet to be fully substantiated, but preliminary clustering results
from the 1\,kpc SPYGLASS analysis (R.~Kerr et al.\ in prep) suggest
the presence of contiguous stellar populations connecting each of
these groups in space-velocity coordinates.  In other words, the stars
appear to be comoving, though with a continuous gradient in velocity
as a function of position\replaced{ (Figure~\ref{fig:XYZvtang}, lower
panels).}{. The lower panels of Figure~\ref{fig:XYZvtang} show this in
detail, where $v_b$ is the distance-corrected proper motion in the
direction of increasing galactic latitude, and $v_{l^*}=v_l \cos b$ is
the distance-corrected proper motion in the direction of increasing
galactic longitude after accounting for the local tangent plane
correction.  Some, but not all, of the gradient in the $v_{l^*}$-$l$
plane can be understood through a projection effect stemming from the
Sun's motion with respect to the local standard of rest (see also
Figure 11 by \citet{Zari2018}).  }
In this work, our primary interest in \replaced{the region}{this
region of sky} stems
from the fact that a portion of it was observed by Kepler
(Figure~\ref{fig:XYZvtang}, top panel).  To further explore this
sub-population, we select candidate Cep-Her members through four
steps, the first three being identical to those described in Section~3
of \citet{Kerr2021}.  We briefly summarize them here.

@@

The CAMDs show that for RSG-5, all but one of the candidate
members are on a tight pre-main-sequence locus.  \added{Quantiatively,
88/89 stars with $\bpmrpo \geq 1.5$ are consistent with being on the
pre-main-sequence.} This implies a false
positive rate of a few percent, at most.  By comparison, our control
sample (the $\delta$~Lyr candidates) has a false positive rate of
$\approx$12\%, based on the number of stars that photometrically
overlap with the field population, while being outliers from the bulk
cluster population.  For CH-2, our membership selection gives 27
objects in the color range displayed, and 23 of them appear to be
consistent with being on the pre-main-sequence.  This implies a false
positive rate in CH-2 of $\approx$15\%.

@@

We selected stars suitable for gyrochronology by requiring 
\replaced{$\bpmrpo \geq 0.5$}{$\bpmrpo \geq 0.6$} to focus on FGKM stars that experience magnetic
braking.
For TESS, we also restricted our sample to $G<16$, to ensure the stars
are bright enough to extract usable light curves from the full-frame
images.  The magnitude cut corresponds to $\bpmrpo < 2.6$
($\sim$M3V) at the relevant distances.  These cuts gave 19 stars in
CH-2 and \replaced{47}{42} stars in RSG-5.  We extracted light curves from the TESS
images using the \texttt{unpopular} package \citep{hattorio_2021_cpm},
and regressed them against systematics with its causal pixel model.
We measured rotation periods using Lomb-Scargle periodograms and
visually vetted the results using an interactive program that allows
us to switch between TESS Cycles, select particular sectors, flag
stars with multiple periods, and correct half-period harmonics. For
ZTF, we used the same color cut to focus on FGKM stars, but restricted
the sample to $13 < G < 18$ to avoid the saturation limit on the
bright end and ensure sufficient photometric precision at the faint
end. We followed the procedure outlined in \citet{curtis_rup147_2020}:
we downloaded $8'\times8'$ image cutouts, ran aperture photometry for
the target and neighboring stars identified with Gaia, and used them
to define a systematics correction to refine the target light curves.

The lower panels of Figure~\ref{fig:age} show the results.  In RSG-5,
\replaced{41/47}{39/42} stars have rotation periods \replaced{faster
than}{at least as fast as} the Pleiades (\replaced{87}{93}\%).  This
numerator omits the two stars with periods $>$$12$\,days visible in
the lower-left panel of Figure~\ref{fig:age}.  The age interpretation
for these latter stars, particularly the $\approx$M2.5 dwarf, is not
obvious.  \citet{rebull_usco_2018} for instance have found numerous
M-dwarfs with 10-12 day rotation periods at ages of USco
($\sim$$8$\,Myr), and some \replaced{may}{do} still exist at ages of
LCC ($\sim$$16$\,Myr; \replaced{L.~Rebull
submitted}{\citealt{2022AJ....164...80R}}).  Regardless, since nearly
no field star outliers seem to be present on the RSG-5 CAMD, the fact
that we do not detect rotation periods for $\approx$\replaced{13}{7}\%
of stars should perhaps be taken as an indication for the fraction of
stars for which rotation periods might not be detectable, due to
{e.g.}, pole-on stars having lower amplitude starspot modulation.

For CH-2, 13/19 stars have rotation periods that are obviously faster
than their counterparts in the Pleiades\deleted{ (68\%)}.  4 stars, not included
in the preceding numerator, are M-dwarfs with rotation periods between
10 and 12.5 days.  As previously noted, the age interpretation for
these M-dwarfs is not obvious.  \added{If none are cluster members,
the rotation period detection fraction is 68\%; if all are members, it
is 89\%}.
\replaced{Regardless, the $\approx$15\% false
positive rate for CH-2 determined from the CAMD seems consistent with
our fraction of detected rotation periods; RSG-5 was missing rotation
periods for $\approx$15\% of its candidate members, even though all of
its stars appear photometrically consistent with being on a single
pre-main-sequence locus.}{
This sets an upper bound on the contaminant fraction in our candidate
CH-2 members at about one in three.  The true value may be as low as 
one in six, based on our earlier CAMD analysis.
}
